Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow

### 2016-08-04 paper/widegap_mri_paper.tex
more widegap intro

The magnetorotational instability (MRI) drives angular momentum transport and turbulence in astrophysical disks. Its discovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}; actually a rediscovery of \citei{Chandrasekhar:1960wh}) was a breakthrough in the longstanding question of how efficient accretion can exist in the universe: that is, how matter collapsing onto a central body is able to coalesce despite the conservation of specific angular momentum. %The ubiquity of disks ...

Magnetic fields were once assumed to have a stabilizing effect

%The MRI is fundamental to our understanding of star formation, but its 
- people make a number of simplifications, incl thin gap, shearing box, etc

>>>>

The magnetorotational instability (MRI) drives angular momentum transport and turbulence in astrophysical disks. Its discovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}; actually a rediscovery of \citei{Chandrasekhar:1960wh}) was a breakthrough in the longstanding question of how efficient accretion can exist in the universe: that is, how matter collapsing onto a central body is able to coalesce despite the conservation of specific angular momentum. %The ubiquity of disks ...
The ubiquity of astrophysical disks suggests that systems both experience and overcome this centrifugal barrier, and accretion proceeds even in hydrodynamically stable disks. The MRI is excited by weak magnetic fields in differentially rotating fluids, and since its discovery it has been widely invoked to explain accretion in protoplanetary disks () and disks around black holes (), as well as

%Magnetic fields were once assumed to have a stabilizing effect

%The MRI is fundamental to our understanding of star formation, but its 
%- people make a number of simplifications, incl thin gap, shearing box, etc
